% easychair.tex,v 3.5 2017/03/15

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

\usepackage{doc}
\usepackage{minted}
%% \usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bussproofs}
\usepackage{cite}

% For ≞ (requires the LuaLaTeX engine)
\usepackage{unicode-math}
\setmathfont{Stix Two Math}

% Use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}
\newcommand{\Theory}{\texttt{Theory}}
\newcommand{\Proof}{\texttt{Proof}}
\newcommand{\Proposition}{\texttt{Proposition}}
\newcommand{\Bool}{\texttt{Bool}}
\newcommand{\arrow}{\to}
\newcommand{\limp}{\Rightarrow}
\newcommand{\True}{\texttt{True}}
\newcommand{\False}{\texttt{False}}
\newcommand{\STV}[2]{<\!#1, #2\!>}
\newcommand{\arrowtype}{\ \ \text{->}\ \ }
\newcommand{\axon}{\text{ax}_1}
\newcommand{\axtw}{\text{ax}_2}
\newcommand{\axth}{\text{ax}_3}
\newcommand{\axmp}{\text{mp}}
\newcommand{\axmpi}{\text{mp}^i}
\newcommand{\blackbird}{B_1}
\newcommand{\idiot}{I}
\newcommand{\av}{\phi}
\newcommand{\bv}{\psi}
\newcommand{\cv}{\chi}
\newcommand{\dv}{\theta}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
%% \title{Estimating the Probability of a Proposition to be Provable
%%   Using Probabilistic Logic Networks}
%% \title{Estimating the Probability of a Conjecture to be a Theorem with
%%   Probabilistic Logic Networks, and How to Take Advantage of it for
%%   Inference Control}
\title{Using Forward Chaining to Go Backward}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%
\author{Nil Geisweiller}

% Institutes for affiliations are also joined by \and,
\institute{
  SingularityNET Foundation,\\
  Zug, Switzerland\\
  \email{nil@singularitynet.io}
}

%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair
\authorrunning{Geisweiller}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair

\titlerunning{Using Forward Chaining to Go Backward}

\begin{document}

\maketitle

%% \begin{abstract}
%% \end{abstract}

\section{Introduction}

If all you have is a forward chainer, is it possible to emulate a
process that is functionally equivalent to backward chaining?
% This is
% the question we attempt to answer in this paper.
Such question stems
from the observation that some rewriting systems, such as
MORK~\cite{NEXT}, perform better on forward chaining than on backward
chaining~\cite{NEXT}.  This is to be contrasted with some other
rewriting systems such as PeTTa~\cite{NEXT}, a Prolog-based rewriting
system, which performs equally well on forward and backward
chaining~\cite{NEXT}.  Since forward chaining in MORK outperforms
forward chaining in PeTTa~\cite{NEXT}, it gives us an incentive to
emulate backward chaining using forward chaining to obtain a
MORK-based backward chainer that outperforms PeTTa.

\section{Forward vs Backward Chaining}

Forward chaining is a reasoning process that begins with axioms and
apply inference rules to produce theorems.  For instance, using the
axioms and inference rules of a fragment of propositional calculus
(borrowed from Metamath~\cite{NEXT}), given with a type theoretic
representation:
\begin{itemize}
\item $\axon : \phi \to (\psi \to \phi)$ $\ \ \ \ \ \ \ \ \ \ \ \ \ \
  \ \ \ \ \ \ \ \ $ $\axtw : ( \phi \to ( \psi \to \chi ) ) \to ( ( \phi \to \psi ) \to ( \phi \to \chi ) )$
\item $\axmp : ( \phi \to \psi ) \arrowtype \phi \arrowtype \psi$
\end{itemize}
where $\text{->}$ is the arrow type and $\to$ is the implication logic
connector, one can build a proof of the identity theorem in a forward
way:
\begin{enumerate}
\item apply $\axmp$ to $\axtw$ and $\axon$ to obtain
  $\ (\axmp\ \axtw\ \axon) : ( \phi \to \psi) \to (\phi \to \phi)$,
\item apply $\axmp$ to step 1 and $\axon$ to obtain
  $\ (\axmp\ (\axmp\ \axtw\ \axon)\ \axon) : \phi \to \phi$.
\end{enumerate}
On the contrary, a backward chainer starts with the proposition to be
proven, and expands the proof backward till it reaches the axioms.
For instance, the process may start with $x? : \phi \to \phi$ where
$x?$ is a hole, then expand $x?$ backward using $\axmp$ to obtain
$(\axmp\ y?\ z?) : \phi \to \phi$, and keep expanding until obtaining
the full proof
$(\axmp\ (\axmp\ \axtw\ \axon)\ \axon) : \phi \to \phi$.  The
advantage of backward chaining is that the search space is constrained
by proofs that can only prove the target proposition, resulting in
much pruning.  The drawback is that it is more sophisticated,
involving full syntactic unification and maintaining a stack of
hypothetical paths.

\section{Emulating Backward Chaining with Forward Chaining}

Given a theory described by axioms and inference rules, the idea is to
invert the inference rules using the \emph{blackbird}
combinator~\cite{NEXT} to obtained an \emph{inverted theory}.
Applying forward chaining on this inverted theory will have the effect
of producing hypothetical truths that preserve their conclusions,
therefor emulating backward expansion.  For instance, given the theory
described earlier, we apply the blackbird combinator, recalled below:
$$\blackbird : (\cv \arrowtype \dv) \arrowtype (\av \arrowtype \bv \arrowtype \cv) \arrowtype \av \arrowtype \bv \arrowtype \dv$$
to the Modus Ponens inference rule to obtain\footnote{The second
  argument of $\blackbird$ is the inference rule to invert, here
  $\axmp : ( \phi \to \psi ) \arrowtype \phi \arrowtype \psi$, thus
  $( \phi \to \psi ) \arrowtype \phi \arrowtype \psi$ must unify up to
  alpha conversion with $\av \arrowtype \bv \arrowtype \cv$, which it
  does.}:
$$\lambda x.((\blackbird\ x)\ \axmp) : (\bv \arrowtype \cv) \arrowtype (\av
\to \bv) \arrowtype \av \arrowtype \cv$$ which can be read as ``given
a hypothetical proof of $\cv$ assuming $\bv$, produce a hypothetical
proof of $\cv$ assuming $\av \to \bv$ and $\av$'', precisely
corresponding to Modus Ponens backward expansion.  For the sake of
simplicity let us rename this inverted inference rule:
$$\axmpi\ \eqdef\ \lambda x.((\blackbird\ x)\ \axmp)$$
We can now use it in a forward way to prove\footnote{Keep in mind that
  variables are scoped within each line, not across lines, and thus
  unification is performed using alpha conversion when necessary.}
the identity theorem shown earlier.  We start from a trivial
hypothetical truth using the \emph{idiot} combinator~\cite{NEXT}:
\begin{enumerate}
\item[1.] $\idiot : (\av \to \av) \arrowtype (\av \to \av)$
\end{enumerate}
then apply $\axmpi$ twice (backward expansion)
\begin{enumerate}
\item[2.]
  $(\axmpi\ \idiot) : (\psi \to (\av \to \av)) \arrowtype \psi
  \arrowtype (\av \to \av)$
\item[3.]
  $(\axmpi\ (\axmpi\ \idiot)) : (\chi \to (\psi \to (\av \to \av)))
  \arrowtype \chi \arrowtype \psi \arrowtype (\av \to \av)$
\end{enumerate}
then apply the result of step 3 to axiom $\axtw$ (reach the second axiom)
\begin{enumerate}
\item[4.]
  $((\axmpi\ (\axmpi\ \idiot))\ \axtw) : (\av \to (\bv \to \av)) \arrowtype (\av \to \bv) \arrowtype (\av \to \av)$
\end{enumerate}
then apply the result of step 4 to axiom $\axon$ twice (reach the first axiom)
\begin{enumerate}
\item[5.]
  $(((\axmpi\ (\axmpi\ \idiot))\ \axtw)\ \axon) : (\av \to \bv) \arrowtype (\av \to \av)$
\item[6.] $((((\axmpi\ (\axmpi\ \idiot))\ \axtw)\ \axon)\ \axon) : (\av \to \av)$
\end{enumerate}
giving us a prove of the identity theorem.  The proof is equivalent to
the one provided earlier $(\axmp\ \axon\ (\axmp\ \axon\ \axtw))$ but
is constructed inside-out.  The reader may notice that this process of
construction emulates a backward chaining depth-first search, and that
the proof encodes a search path along the search tree.  It is also
possible to emulate breadth-first search by using the \emph{bluebird}
combinator and may be the subject of future work.

\section{Experimental Results}

Experiments have been conducted to compare such backward chaining
emulation with regular backward chaining based on a corpus ported from
Metamath~\cite{NEXT} to MeTTa~\cite{NEXT}.  To establish a fair
comparison, care has been put to ensure that both chainers search
isomorphic spaces and run on the same back-end, PeTTa.
% , to measure the
% overhead of the emulation itself and nothing else.
It has been found that the emulation is twice slower in the worse
cases to almost as fast in the best cases compared to regular backward
chaining~\cite{NEXT}.  This indicates that it could be a competitive
approach on rewriting systems that excel at forward chaining such as
MORK, though more work is required to determine exactly how much.
Although we understand that brute force search, however fast, is never
going to replace inference control~\cite{NEXT}, we believe it still
serves as a base that may underlie the performance of the whole
system, and thus is worth our attention.

\newpage

%------------------------------------------------------------------------------

\bibliographystyle{plain}
\label{sect:bib}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{BackwardViaForward}

\newpage

%------------------------------------------------------------------------------

\appendix

%------------------------------------------------------------------------------
\end{document}
